Renewal and stability in populations structured by remaining years of life

### 2014-09-12 ThanoRepro/ThanoRepro.tex
added justifications, per josh's suggestion

Temporal perspectives and rescalings of the lifecourse are abundant in
demography already. It is my view that these efforts have enriched the
discipline by producing new insights and questions much more than they have obscured it by
producing conundrums. Likely some demographic phenomena are best
described as a function of time since birth, and others of time until death,
while still others can be a function of both to some degree. Fertility and
reproduction are plausibly a function of both age perspectives, and the question
of how to best partition such forces and events over time is a new conundrum
that I do not tackle here. Instead I describe some of the formal consequences
of the Lotka-Euler renewal model when viewed from the thanatological perspective
of age. The renewal model presented here is not necessarily a suggestion for a
better way to model or project population, but in the minimal sense it
represents an alternative way to tell the story of population renewal.

>>>>

Temporal perspectives and rescalings of the lifecourse are abundant in
demography already. It is my view that these efforts have enriched the
discipline by producing new insights and questions much more than they have obscured it by
producing conundrums. Likely some demographic phenomena are best
described as a function of time since birth, and others of time until death,
while still others can be a function of both to some degree. Fertility and
reproduction are plausibly a function of both age perspectives, and the question
of how to best partition such forces and events over time is a new conundrum
that I do not tackle here. Instead I describe some of the formal consequences
of the Lotka-Euler renewal model when viewed from the thanatological perspective
of age.

The renewal model is the core of modern demography, as it elegantly relates
fertility, mortality, and population structure, revealing the long
term consequences of a particular set of conditions. Some might think that
stable population theory is not relevant to everyday demography, since
populations seem to always be changing. Still, most projections seem to trail
off into a stable state, and stability or even stationarity are indeed
imaginable for the world population of humans one day. I present a renewal model
based on thanatological age not necessarily as a suggestion for a
better way to model or project population (or certain aspects of it), although I
plant this is an open and unexplored empirical question.
In a minimal sense the thanatological renewal model represents an alternative
and coherent way to describe the process of population renewal. In a maximal sense, the model may
actually prove useful, as a model relating renewal to a death flow rather than
to a birth flow. This enriches our understanding of lifespan and vitality, but also
of the side effects of our species' renewal, such as the human biomass cycle.

Further, there is likely a range of lifecourse transitions that covary
more with remaining lifetime that with life lived, but these are mostly unstudied at this
time,\footnote{I suppose that the prime candidates are late-life transitions.}
and it is beneficial to expand the formal demographic base by incorporating such information into models. In an even more general sense, by
thinking about remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area
applications, such as education, partering, the economic lifecycle, or migration.
The present inquiry is justified possibly for its own sake, and definitely as an
attempt at accounting for remaining time structure in general in modeling that may prove useful.
